Elastic Scattering in General Relativistic Ray Tracing for Neutrinos

### 2018-06-15 paper.tex
Finish adding Evan's responses to TODO notes (email Feb 28, 2018)

We also solve for the flavor evolution of this system,
integrating $S$ according to Eqn.~\ref{eqn:schroedinger},
as described in \cite{zhu2016-mnr_nsns_remnant}.
We show the survival probabilities for $\nu_e$ and $\bar{\nu}_e$
for the \emph{noscat} treatment in Fig.~\ref{fig:survival-noscat},
and for the \emph{scat} treatment in Fig.~\ref{fig:survival-scat}.
The survival probability is the probability that a neutrino, if measured,
will be found to be in its original flavor state.
The survival probabilities are computed at each point along the trajectory
from the absolute square of the diagonal terms of the flavor evolution matrix,
$P_{\nu_\alpha \to \nu_\alpha} = |S_{\alpha \alpha}|^2$.
When $P_{\nu_e \to \nu_e}$ decreases,
as can be seen for example in Fig.~\ref{fig:survival-scat} for $r>400$~km,
some of the electron-type neutrinos have oscillated into
mu- or tau-type neutrinos.

@@

In the special case of radiative equilibrium we know that the source
term vanishes: an equal number of neutrinos are emitted from
and absorbed by the matter for any length traversed.
We also know in this case that the neutrino distribution function must be
$f^{\rm eq}$, the equilibrium Fermi-Dirac distribution function of
Eqn.~\ref{eqn:feq}.
With these facts we can rerrange Eqn.~\ref{eqn:sources_ae_1} to give us
Kirchoff's Law:
\begin{align}
  \label{eqn:kirchoffs_law}
  j_i &= \frac{\chi_{a,i}}{1-f^{\rm eq}}f^{\rm eq}, \\
  \label{eqn:stimulated_opacity}
  &= \chi^*_{a,i}\,f^{\rm eq},
\end{align}
where in Eqn.~\ref{eqn:stimulated_opacity} we have introduced the opacity
corrected for stimulated absorption, $\chi^*_{a,i}$.

@@

\subsection{Inelastic scattering}
\label{ssec:sources_si}
Neutrino scattering off of electrons is inelastic, changing the
magnitude and direction of the neutrino's momentum.
\footnote{TODO (Evan): citation? maybe \cite{keil2003-pinched_spectra}?}
A similar formalism to the above could be used to derive source terms
for inelastic scattering.
\footnote{TODO (Evan): can we compare relative amounts of inelastic to
  elastic scattering?}
We save that for future work, pointing out here that inelastic scattering
treated this way is very sensitive to the energy-dependence of
the background field.

\subsection{Thermal pair annihilation and production}
\label{ssec:sources_pp}
We do not include thermal pair processes within the above formalism,
\footnote{TODO (Evan): can we explain this choice better? and add citation?}
but as mentioned in App.~\ref{ssec:sources_ae}, we do include
these processes in an effective absorption opacity for $\mu$ and $\tau$
neutrinos and antineutrinos. \lstinline{NuLib} follows
\cite{brue1985-core_collapse,burr2006-neutrino_opacities}
by computing the emissivity of the pair processes in
\footnote{TODO (Evan): improve description?}
Tab.~\ref{tab:neutrino_processes}, and applying Kirchoff's law,
Eqn.~\ref{eqn:kirchoffs_law}, to convert this to an effective absorption opacity,
and its corresponding emissivity obeying Kirchoff's law.
We only use this effective opacity for heavy-lepton neutrinos because
it is subdominant to the true absorption opacity for
$\nu_e$ and $\bar{\nu}_e$.

>>>>

We also solve for the flavor evolution of this system,
integrating $S$ according to Eqn.~\ref{eqn:schroedinger},
as described in \cite{zhu2016-mnr_nsns_remnant}.
We show the survival probabilities for $\nu_e$ and $\bar{\nu}_e$
for the \emph{noscat} treatment in Fig.~\ref{fig:survival-noscat},
and for the \emph{scat} treatment in Fig.~\ref{fig:survival-scat}.
The survival probability is the probability that a neutrino, if measured,
will be found to be in its original flavor state.
The survival probabilities are computed at each point along the trajectory
from the absolute square of the diagonal terms of the flavor evolution matrix,
$P_{\nu_\alpha \to \nu_\alpha} = |S_{\alpha \alpha}|^2$.
When $P_{\nu_e \to \nu_e}$ decreases,
as can be seen for example in Fig.~\ref{fig:survival-scat} for $r>400$~km,
some of the $e$ neutrinos have oscillated into $\mu$ or $\tau$ neutrinos.

@@

In the special case of radiative equilibrium we know that the source
term vanishes: an equal number of neutrinos are emitted from
and absorbed by the matter for any length traversed.
We also know in this case that the neutrino distribution function must be
$f^{\rm eq}$, the equilibrium Fermi-Dirac distribution function of
Eqn.~\ref{eqn:feq}.
With these facts we can rearrange Eqn.~\ref{eqn:sources_ae_1} to give us
Kirchoff's Law:
\begin{align}
  \label{eqn:kirchoffs_law}
  j_i &= \frac{\chi_{a,i}}{1-f^{\rm eq}}f^{\rm eq}, \\
  \label{eqn:stimulated_opacity}
  &= \chi^*_{a,i}\,f^{\rm eq},
\end{align}
where in Eqn.~\ref{eqn:stimulated_opacity} we have introduced the opacity
corrected for stimulated absorption, $\chi^*_{a,i}$.

@@

\subsection{Inelastic scattering}
\label{ssec:sources_si}
Neutrino scattering off of electrons is inelastic, changing the
magnitude and direction of the neutrino's momentum \cite{brue1985-core_collapse}.
In a supernova environment we expect inelastic scattering off of electrons
and nucleons to shift the neutrino spectra to lower energies, most noticeably
for heavy-lepton neutrinos \cite{thom2003-ccsne_neutrinos}.
A similar formalism to the above could be used to derive source terms
for inelastic scattering.
We save that for future work, pointing out here that inelastic scattering
treated this way is very sensitive to the energy-dependence of
the background field. In this work we take $C_{\rm SI}=0$.

\subsection{Thermal pair annihilation and production}
\label{ssec:sources_pp}
We do not include thermal pair processes within the standard pair-process
formalism; in other words we take $C_{\rm PP}=0$.
But as mentioned in App.~\ref{ssec:sources_ae}, we do include
these processes in an effective emission/absorption opacity for $\mu$
and $\tau$ neutrinos and antineutrinos.
This has been shown to work well for core-collapse supernovae
\cite{ocon2015-gr1d_with_nu}. Within \lstinline{NuLib}, we compute
the energy-dependent emissivity of the pair processes in
Tab.~\ref{tab:neutrino_processes}, ignoring final state neutrino blocking.
We then apply Kirchoff's law, Eqn.~\ref{eqn:kirchoffs_law},
to convert this to an effective absorption opacity.
We only use this effective opacity for heavy-lepton neutrinos
because the charged-current emission/absorption processes for $\nu_e$
and $\bar{\nu}_e$ dominate in the environments we study: early merger
remnants and supernovae.
